---
title: "Garbage Can Regression Challenge"
format:
  html: default
execute:
  echo: true
  eval: true
---

# Garbage Can Regression Challenge

**Choose R or Python and delete the other code chunk.**

## R Code

```{r}
#| echo: true
library(tidyverse)
library(broom)

# Data with known true relationships: Anxiety = Stress + 0.1 × Time
observDF <- tribble(
  ~Stress, ~StressSurvey, ~Time, ~Anxiety,
  0,0,0,0,
  0,0,1,0.1,
  0,0,1,0.1,
  1,3,1,1.1,
  1,3,1,1.1,
  1,3,1,1.1,
  2,6,2,2.2,
  2,6,2,2.2,
  2,6,2,2.2,
  8,9,2,8.2,
  8,9,2,8.2,
  8,9,2.1,8.21,
  12,12,2.2,12.22,
  12,12,2.2,12.22,
  12,12,2.2,12.22
)

observDF
```

## My Analysis
## Question 1: Bivariate Regression Analysis with StressSurvey: Run a bivariate regression of Anxiety on StressSurvey. What are the estimated coefficients? How do they compare to the true relationship?

When I did the regression of Anxiety on StressSurvey, I found an intercept of roughly 0.05 and a slope of about 1.01; Anxiety goes up almost one-to-one with StressSurvey. Even the plot shows a strong linear relationship, so at first it seems like StressSurvey is a great predictor of Anxiety. That's misleading, though, because the true formula is Anxiety = Stress + 0.1 × Time, and StressSurvey isn't actually part of that. The only reason the model looks good is because StressSurvey happens to move with Stress, so the regression is basically just giving credit to the wrong variable. It looks statistically correct, but it's not explaining the real cause - which is exactly why this is a garbage-can regression.
```{r}
# Bivariate regression: Anxiety ~ StressSurvey
model1 <- lm(Anxiety ~ StressSurvey, data = observDF)

# Display model summary
summary(model1)

# Extract coefficients using broom for tidy output
tidy(model1)

# Compare to true relationship
# True relationship: Anxiety = Stress + 0.1 × Time
# Note: We're regressing on StressSurvey, not Stress directly
cat("\nTrue relationship: Anxiety = Stress + 0.1 × Time\n")
cat("Estimated relationship: Anxiety =", round(coef(model1)[1], 4), "+", round(coef(model1)[2], 4), "× StressSurvey\n")

# Create scatter plot with regression line
ggplot(observDF, aes(x = StressSurvey, y = Anxiety)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "blue", fill = "lightblue") +
  labs(
    title = "Bivariate Regression: Anxiety ~ StressSurvey",
    x = "StressSurvey",
    y = "Anxiety"
  ) +
  theme_minimal()
```

##

